---
title: "Bighorn baro compare"
author: "Megan Sears"
date: "2024-11-8"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
editor_options: 
  chunk_output_type: console
---

```{r open packages, include=FALSE}

library(tidyverse)
library(plotly)
library(RColorBrewer)
library(hexbin)

theme_set(theme_bw(base_size = 20))

```

```{r, include = F}

# prep data
baro1 <- read_csv('./raw_data/bighorn_baro_composite_new.csv') %>%
  mutate(baro = 'original',
         datetime = mdy_hm(datetime)) %>%
  filter(datetime > ymd('2024-06-04'))

baro2 <- read_csv('./raw_data/bighorn_baro2_composite.csv') %>%
  mutate(baro = 'shade',
         datetime = mdy_hm(datetime))

baro_combine <- bind_rows(baro1, baro2)

```

# Baro time series

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6}

compare <- ggplot(baro_combine, 
                  aes(x = datetime, 
                      y = P_kPa, 
                      color = baro)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2",
                     direction=-1) 

ggplotly(compare)

```

# 1:1 comparison

Baro shade vs baro original

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6}
# left join so can compare 1:1
baro_1to1 <- left_join(baro2, baro1, by = 'datetime')

# get r-square of the 1:1
lm <- lm(P_kPa.y ~ P_kPa.x, 
         data = baro_1to1)

r_2 <- summary(lm)$r.squared

oneone <- ggplot(baro_1to1, aes(P_kPa.x, 
                      P_kPa.y)) + 
  geom_point() +
  geom_abline(slope = 1, 
              intercept = 0, 
              color = "red", 
              linetype = "dashed") +
  labs(y = 'Original', 
       x = 'Shade') +
  annotate("text", 
         x = -Inf, 
         y = Inf, 
         hjust = -0.3, 
         vjust = 2, 
         label = paste("R² =", round(r_2,2)), 
         color = "blue")


ggplotly(oneone)

```

# Baro difference time series

- Shade minus original = difference
- If it's negative: original baro pressure > shade baro pressure
- If it's positive: original < shade

```{r, echo = FALSE, message = FALSE, warning = FALSE,fig.width = 10, fig.height = 6}

# Create the difference column
baro_1to1 <- baro_1to1 %>%
  mutate(difference = P_kPa.x - P_kPa.y,
         hour = hour(datetime) + minute(datetime) / 60)

ts_diff <- ggplot(baro_1to1, aes(x = datetime, 
                      y = difference)) +
  geom_point() +
  labs(x = 'Datetime', 
       y = 'Difference (P_kPa)')

ggplotly(ts_diff)

```

# Baro time of day difference 

- Shade minus original = difference
- If it's negative: original baro pressure > shade baro pressure
- If it's positive: original < shade

```{r, echo = FALSE, message = FALSE, warning = FALSE,fig.width = 10, fig.height = 6}

# Plot with time of day on x-axis and difference on y-axis
tod <- ggplot(baro_1to1, aes(x = hour, 
                      y = difference)) +
  geom_point() +
  labs(x = 'Time of Day', 
       y = 'Difference (P_kPa)')

ggplotly(tod)

# hex bins to see counts
tod_hex <- ggplot(baro_1to1, aes(x = hour, y = difference)) +
  geom_hex(bins = 40) + 
  scale_fill_viridis_c() +
  labs(x = 'Time of Day', y = 'Difference (P_kPa)')

ggplotly(tod_hex)

```

# What does this mean for stage?

Stage using shade baro and original baro

```{r, echo = FALSE, message = FALSE, warning = FALSE,fig.width = 10, fig.height = 6}

stage_pt <- read_csv('./raw_data/bighorn_stage_composite_new.csv') %>%
  rename(Pw_kPa = 2,
         datetime = 1) %>%
  mutate(datetime = mdy_hm(datetime))

combine <- left_join(baro_1to1,
                     stage_pt,
                     by = 'datetime') %>%
  mutate(stage_shade_kPa = Pw_kPa - P_kPa.x,
         stage_orig_kPa = Pw_kPa - P_kPa.y) %>%
  mutate(stage_shade_cm = (stage_shade_kPa*101.9716)/10,
         stage_orig_cm = (stage_orig_kPa*101.9716)/10) 
         
# Extract specific colors from Dark2
dark2_colors <- brewer.pal(n = 8, name = "Dark2")  # Extracts 8 colors from Dark2
orange_color <- dark2_colors[2]
green_color <- dark2_colors[1]

stage_TS <- ggplot(combine, aes(x = datetime, 
                    y = stage_orig_cm, 
                    color = "Original")) +
  geom_line() +
  geom_line(aes(y = stage_shade_cm, 
                color = "Shade")) +
  scale_color_manual(values = c("Original" = orange_color, 
                                "Shade" = green_color)) +
  labs(x = 'Datetime', y = 'Stage_cm') 

ggplotly(stage_TS)

```

# Difference in stage

- Shade minus original = difference
- If it's negative: original stage > shade stage
- If it's positive: original < shade

```{r, echo = FALSE, message = FALSE, warning = FALSE,fig.width = 10, fig.height = 6}

combine <- combine %>%
  mutate(stage_diff = stage_shade_cm - stage_orig_cm)

stage_TS_diff <- ggplot(combine, aes(x = datetime, 
                    y = stage_diff)) +
  geom_line() +
  labs(x = 'Datetime', y = 'Stage difference (cm)') 

ggplotly(stage_TS_diff)

```

